---
title: "General distance & error based"
output: pdf_document
date: "2025-03-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# General
```{r}
library(sdcSpatial)
```

## Defining maps and plotting
Here using the SDC technique smoothing.
And assuming that any NA values amount to 0.
```{r}
# the maps
map_original = sdc_raster(dwellings[,1:2], dwellings$consumption)
map_safe = protect_smooth(map_original)

# the plots
mfrow = c(1,2)
plot(map_original)
plot(map_safe)

# the data
## turning raster into matrices
dat_original = raster::as.matrix(map_original$value$mean); dat_original[is.na(dat_original)] = 0
dat_safe = raster::as.matrix(map_safe$value$mean); dat_safe[is.na(dat_safe)] = 0
```

# Utility measures (Global)
## MSE
MSE is the mean squared error, computed between the original and the perturbed dataset.
For each predefined grid, take the mean of the squared difference between the original and the safe map
```{r}
MSE = mean((dat_original - dat_safe)^2)
MSE

MSE_func = function(original, safe){
  mean((original - safe)^2)
}
```

## MAE
```{r}
MAE = mean(abs(dat_original - dat_safe))
MAE
```

## SME
Have to make a decision on what to do with zeros, could also omit and adjust the length(dat_original) to length(delta).
Has an upper bound of 2 instead of one, could also omit the *2
Not symmetric wrt too high or too low values, if original is 10 and safe is 5 then delta 5/15, but if safe is 15 then delta 5/25.
```{r}
delta = abs(dat_original - dat_safe)/(dat_original + dat_safe)
delta[is.nan(delta)] = 0
SME = 1/(length(dat_original)*2) * sum(delta)
SME
```

## Hellinger Distance
Between 0 and sqrt(N), where N is the sum of counts over the whole map.
```{r}
in.sqrt = sum((sqrt(dat_safe) - sqrt(dat_original))^2)
HD = 1/sqrt(2) * sqrt(in.sqrt)
HD
# percent of maximum
HD.max = sqrt(length(dat_safe))
HD/HD.max
```

## RAD
Again the problems with the zeros, but according to Young (2008) source, set to 0 if og is 0
```{r}
RAD.int = abs(dat_original - dat_safe)/dat_original
RAD.int[is.nan(RAD.int)] = 0
RAD = mean(RAD.int)
RAD
```

## RMSE
```{r}
RMSE = sqrt(MSE)
RMSE
```

# Utility measures (local)
Error for each cell, plotted as maps using the function below
```{r}
local_error_map = function(data){
  df_long = reshape2::melt(data)
  
  # Generate a blue color palette
  blue_palette <- colorRampPalette(c("white", "blue", "darkblue"))
  
  # Normalize values to a range from 1 to the number of colors
  num_colors <- 1000 # Adjust for more color granularity
  col_breaks <- cut(df_long$value, breaks = num_colors, labels = FALSE)
  
  # Assign colors based on values
  point_colors <- blue_palette(num_colors)[col_breaks]
  
  ## plotting the final map
  plot(df_long$Var1, df_long$Var2, col = point_colors, pch = 19, asp = 1,
  main = "Error per cell")
}

local_error_map(delta)

```
