---
title: "hospot_shape"
output: pdf_document
date: "2025-04-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## general
```{r}
# retrieve basic functions for contours and center/scaling
source("hotspot_functions.R")

## libraries
library(dplyr)
library(ggplot2)
library(lidaRtRee)
library(sf)
library(purrr)
library(sdcSpatial)
library("tigers")
library(sp)
library(raster)

## data
set.seed(1)
df.original <- base::data.frame(x = c(abs(rnorm(200)), abs(rnorm(200,4))), y = c(abs(rnorm(200)), abs(rnorm(200,4))))
df.safe = mask_random(df.original, r = 0.5)

plot(df.original);points(df.safe, col = "red")
```

Now using the unbounded hotspots function
```{r}
og.hs = unbounded_hotspots(df.original, minimum = 0.05)
sf.hs = unbounded_hotspots(df.safe, minimum = 0.05)

plot(og.hs$geometry, axes = TRUE, col = "blue")              # the original hotspots
plot(sf.hs$geometry, axes = TRUE, add = TRUE, col = "green") # the safe hotspots

```
Generally we will compare hotspots (safe and og) based on their proximity to each other. How to do this will come later
For now take simply the big hotspots to explore the measures first
```{r}
og.hs1 = og.hs[1,1]
sf.hs1 = sf.hs[1,1]


plot(og.hs1$geometry, axes = TRUE, col = "blue")              # the original hotspots
plot(sf.hs1$geometry, axes = TRUE, add = TRUE, col = "green") # the safe hotspots
```

## symmetric difference
```{r}
sc.og.hs = scale.center.poly(og.hs1$geometry)
sc.sf.hs = scale.center.poly(sf.hs1$geometry)
symdiff1 = st_sym_difference(sc.og.hs, sc.sf.hs)

plot(sc.og.hs, axes = TRUE, col = "blue")              # the original hotspots
plot(sc.sf.hs, axes = TRUE, add = TRUE, col = "green") # the safe hotspots
plot(symdiff1, add = TRUE, col = "red")

# the to calculate the area of the symmetric difference
st_area(symdiff1)
```

## Hausdorff distance
https://rdrr.io/cran/tigers/man/HausdorffDistance.html
```{r}
library("tigers")

points.og1 = st_cast(sc.og.hs, "MULTIPOINT") %>% as("Spatial") %>% coordinates()
points.sf1 = st_cast(sc.sf.hs, "MULTIPOINT") %>% as("Spatial") %>% coordinates()

# does not give the answer
HausdorffDistance(points.og1, points.sf1, directed = FALSE)

```
FINAL VERSION
we know this works at least in the right way and it is still widely available so going with this Hausdorff distance from tigers
```{r}
A = rbind(c(0,0), c(4,0), c(0,2))
B = rbind(c(2,2), c(6,4), c(2,4))

HausdorffDistance <- function(A, B, directed = FALSE)
{
    s <- 1:nrow(A)
    D <- as.matrix(dist(rbind(A, B)))
    D <- D[s, -s]
    h <- max(apply(D, 1, min))
    if (directed) return(h)
    max(h, apply(D, 2, min))
}

HausdorffDistance(A,B)
```


## Frechet distance
Based on an existing function, but we have two things to take into account:
- we need to pick as starting points the point closest to each other (as specified), could be solved by changing the input
- going clockwise from both point sets, also be solved by changing the input

```{r}
## first we calculate the starting points
###### first starting point
first_starting_point = function(C, D, first = c(0,0)){
    
  all_points = as.data.frame(cbind(rbind(C,D), group = as.factor(c(rep("C", nrow(C)), rep("D", nrow(D)))))) # include and indicator from which polygon they are
  all_points = all_points[!duplicated(all_points[,1:2]),]
    
  # find distance to origin 
  all_points$dist2origin = sqrt((all_points[,1] - first[1])^2 + (all_points[,2] - first[2])^2)
  
  # find the points that are closest to the origin (could be multiple from the same group and from different groups)
  closest2origin = all_points %>% filter(dist2origin == min(dist2origin)) 

  # if we have more than one row we calculate the minimal distance of those points to the other group
  if (nrow(closest2origin) == 1) start1 = closest2origin 
  else {
    closest2origin$dist2other = apply(closest2origin, 1, function(x) {
      other_group = all_points[all_points$group != as.numeric(x[3]),] # define other group
      min(sqrt((x[1] - other_group[,1])^2 + (x[2]- other_group[,2])^2))}) # minimal distance to other group
    
    closest2other = closest2origin %>% filter(dist2other == min(dist2other)) 

    if (nrow(closest2other) == 1) start1 = closest2other 
    else {
      smallest.x = closest2other[which(closest2other$V1 == min(closest2other$V1)), ]
      
      if(nrow(smallest.x) == 1) start1 = smallest.x 
      else start1 = closest2other[which(smallest.x$V2 == min(smallest.x$V2)), ]
    }
  }
  
  return(start1)
  
}

##### second starting point
second_starting_point = function(C,D,start) {
  
  all_points = as.data.frame(cbind(rbind(C,D), group = as.factor(c(rep("C", nrow(C)), rep("D", nrow(D)))))) # include and indicator from which polygon they are
  all_points = all_points[!duplicated(all_points[,1:2]),]
  relevant_points = all_points %>% filter(group != start$group)

  # if the first starting point also occurs in the other group this is automatically the second starting point
  if(nrow(merge(data.frame(start[,1:2]), relevant_points[,1:2]))>0) start2 = start
  else {
    relevant_points$dist2start = sqrt((relevant_points[,1] - start[1,1])^2 + (relevant_points[,2] - start[1,2])^2)
    closest2start = relevant_points %>% filter(dist2start == min(dist2start)) 
    
    if (nrow(closest2start) == 1) start2 = closest2start 
    else {
      closest2start$dist2other = apply(closest2start, 1, function(x) {
        other_group = all_points[all_points$group != as.numeric(x[3]),] # define other group
        min(sqrt((x[1] - other_group[,1])^2 + (x[2]- other_group[,2])^2))}) # minimal distance to other group
      
      closest2other = closest2start %>% filter(dist2other == min(dist2other)) 
  
      if (nrow(closest2other) == 1) start2 = closest2other 
      else {
        smallest.x = closest2other[which(closest2other$V1 == min(closest2other$V1)), ]
        
        if(nrow(smallest.x) == 1) start2 = smallest.x 
        else start2 = closest2other[which(smallest.x$V2 == min(smallest.x$V2)), ]
      }
    }    
    
  }
  
  return(start2)
}


#### rotate to clockwise if necessary
is_clockwise <- function(x, y) {
  n <- length(x)
  x_next <- c(x[-1], x[1])
  y_next <- c(y[-1], y[1])
  area <- sum(x * y_next - x_next * y)
  return(area < 0)
}

make_clockwise = function(coordinates){
  # check direction and change if needed
  direction = is_clockwise(coordinates[,1], coordinates[,2])
  
  if(!direction) return(coordinates[nrow(coordinates):1, ])
  else return(coordinates)
  
}


### actual calculating of the frechet distance
compute_frechet_matrix <- function(P, Q) {
  m <- nrow(P)
  n <- nrow(Q)
  D <- matrix(0, nrow = m, ncol = n)
  
  for (i in 1:m) {
    for (j in 1:n) {
      d <- euclidean(P[i, ], Q[j, ])
      if (i == 1 && j == 1) {
        D[i, j] <- d
      } else if (i == 1) {
        D[i, j] <- max(D[i, j - 1], d)
      } else if (j == 1) {
        D[i, j] <- max(D[i - 1, j], d)
      } else {
        D[i, j] <- max(min(D[i - 1, j], D[i - 1, j - 1], D[i, j - 1]), d)
      }
    }
  }
  
  return(D)
  
}
```

```{r}
C <- matrix(c((c(-1,-2,1.5,2,-1) %>% rev()),(c(-1,2,1.5,-2,-1) %>% rev())), ncol = 2, byrow = FALSE)
D <- matrix(2*c(-0.5,-0.75,-2,.5,2,-.5,
                -0.75,-0.5,2,.5,-2,-.75), ncol = 2, byrow = FALSE)

calc_frechet_distance = function(C,D){
  # remove duplicates
  C = C[!duplicated(C),]
  D = D[!duplicated(D),]
  
  firsths = first_starting_point(C,D)
  secondhs = second_starting_point(C,D,firsths)
  
  Cstart = which(C[,1] == firsths[,1] & C[,2] == firsths[,2])
  Dstart = which(D[,1] == secondhs[,1] & D[,2] == secondhs[,2])
  
  Corder = rbind(C[Cstart:nrow(C),], C[1:Cstart,])
  Dorder = rbind(D[Dstart:nrow(D),], D[1:Dstart,])
  
  Crotate = make_clockwise(Corder)
  Drotate = make_clockwise(Dorder)
  
  fm = compute_frechet_matrix(Crotate, Drotate)
  frechet_distance <- fm[nrow(Crotate), nrow(Drotate)]
  
  return(frechet_distance)
}

calc_frechet_distance(C,D)

# if rotate/mirror first then frechet existing function can be used
#C = st_polygon(list(Crotate))
#D = st_polygon(list(Drotate))
#st_distance(C,D, which= "Frechet")

```


## box-counting dimension
```{r}
box_count_func = function(eps, poly) {

  bb33 = st_bbox(scale.center.poly(poly))
  bbsize = max(abs(bb33 - 0))
  r <- raster(nrows=1*eps, ncol=1*eps, xmn=-bbsize, ymn=-bbsize, xmx = bbsize, ymx = bbsize)
  vals <- seq(1:length(r))
  
  r2 = st_as_sf(rasterToPolygons(r))

  inter2 = st_overlaps(poly, sf::st_set_crs(r2, st_crs(poly)), byid = TRUE)
  ind = inter2[[1]]
  
  
  values(r) = NA
  values(r)[ind] = vals[ind]
  
  plot(r)
  plot(poly, add = TRUE)

  return(list("boxes" = length(ind),
              "size" = 2*bbsize / eps,
              "eps" = eps))
  
  
}

bc1 = box_count_func(5,scale.center.poly(og.hs[1,]$geometry))


```

```{r}
bc2 = sapply(1:100, function(x) box_count_func(x,scale.center.poly(og.hs[1,]$geometry)))



```
```{r}
bc3 = sapply(1:100, function(x) box_count_func(x,scale.center.poly(sf.hs[1,]$geometry)))
```


```{r}
num = sapply(bc2[1,], log) 
denom = sapply(bc2[3,], function(x) log(1/x)) 
bcdim = num/denom

num2 = sapply(bc3[1,], log) 
denom2 = sapply(bc3[3,], function(x) log(1/x)) 
bcdim2 = num2/denom2

plot(as.vector(bc2[3,]),bcdim, type = "l", col = "blue")
lines(as.vector(bc3[3,]),bcdim2, col = "red")

```


### Angle measure (ONLY FOR GRIDS!!!!)
AND NO SCALING!!!
```{r}

map_original = sdc_raster(dwellings[,1:2], dwellings$consumption)
map_safe = protect_smooth(map_original)

dat_original = raster::as.matrix(map_original$value$mean); dat_original[is.na(dat_original)] = 0
dat_safe = raster::as.matrix(map_safe$value$mean); dat_safe[is.na(dat_safe)] = 0


gridsf = grid_hotspot(dat_safe, 5000)
gridog = grid_hotspot(dat_original, 5000)


gridsf2 = st_as_sf(gridsf[1,])
gridog2 = st_as_sf(gridog[1,])

sgrid2 = scale.center.poly(gridsf2$geometry)
ogrid2 = scale.center.poly(gridog2$geometry)

plot(ogrid2, axes = TRUE, col = "blue")              # the original hotspots
plot(sgrid2, axes = TRUE, add = TRUE, col = "green") # the safe hotspots


plot(gridog2, axes = TRUE, col = "blue")              # the safe hotspots
plot(gridog2, axes = TRUE, add = TRUE, col = "green") # the og hotspots

```

```{r}
number_angels = function(hotspot){
  ogco = st_coordinates(st_cast(hotspot, "POLYGON"))
  changedf = data.frame(
  x = c(diff(ogco[,1]),diff(ogco[,1])[1]),
  y = c(diff(ogco[,2]),diff(ogco[,2])[1]))


  sum(diff(changedf$x) != 0 & diff(changedf$y) != 0)
}


angle_measure = function(safe, original){
  area_safe = st_area(safe)
  area_og = st_area(original)
  
  angles_safe = number_angels(safe)
  angles_og = number_angels(original)
    
  (angles_safe * area_og) / (angles_og * area_safe)
}


angle_measure(gridsf2, gridog2)


```


##### Correlation dimension
```{r}
corr_integral = function(x, eps){
  N = nrow(x)
  
  heavyside = sum((eps - dist(x)) > 0)
  
  (1/(N*(N-1))) * heavyside
  
}


corr_dimension_func = function(n, polygon, step = 0.01){
  sampledpoly = st_sample(polygon, n)
  sampled = st_cast(sampledpoly, "MULTIPOINT") %>% as("Spatial") %>% coordinates()
  epsis = seq(0,max(dist(sampled)), step)
  Ce = sapply(epsis, function(eps) corr_integral(sampled,eps))

  return(list(
    "eps" = epsis,
    "C(e)" = Ce
  ))
}

ogcorr = corr_dimension_func(300,sc.og.hs)
sfcorr = corr_dimension_func(300,sc.sf.hs)

plot(log(ogcorr$eps), log(ogcorr$`C(e)`), type = "l", col = "blue")
lines(log(sfcorr$eps), log(sfcorr$`C(e)`), type = "l", col = "red")
```




























