---
title: "hospot_shape"
output: pdf_document
date: "2025-04-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## general
```{r}
# retrieve basic functions for contours and center/scaling
source("hotspot_functions.R")

## libraries
library(dplyr)
library(ggplot2)
library(lidaRtRee)
library(sf)
library(purrr)
library(sdcSpatial)
library("tigers")
library(plyr)


## data
set.seed(1)
df.original <- data.frame(x = c(abs(rnorm(200)), abs(rnorm(200,4))), y = c(abs(rnorm(200)), abs(rnorm(200,4))))
df.safe = mask_random(df.original, r = 0.5)

plot(df.original);points(df.safe, col = "red")
```

Now using the unbounded hotspots function
```{r}
og.hs = unbounded_hotspots(df.original, minimum = 0.05)
sf.hs = unbounded_hotspots(df.safe, minimum = 0.05)

plot(og.hs$geometry, axes = TRUE, col = "blue")              # the original hotspots
plot(sf.hs$geometry, axes = TRUE, add = TRUE, col = "green") # the safe hotspots
```
Generally we will compare hotspots (safe and og) based on their proximity to each other. How to do this will come later
For now take simply the big hotspots to explore the measures first
```{r}
og.hs1 = og.hs[1,1]
sf.hs1 = sf.hs[1,1]


plot(og.hs1$geometry, axes = TRUE, col = "blue")              # the original hotspots
plot(sf.hs1$geometry, axes = TRUE, add = TRUE, col = "green") # the safe hotspots
```

## symmetric difference
```{r}
sc.og.hs = scale.center.poly(og.hs1$geometry)
sc.sf.hs = scale.center.poly(sf.hs1$geometry)
symdiff1 = st_sym_difference(sc.og.hs, sc.sf.hs)

plot(sc.og.hs, axes = TRUE, col = "blue")              # the original hotspots
plot(sc.sf.hs, axes = TRUE, add = TRUE, col = "green") # the safe hotspots
plot(symdiff1, add = TRUE, col = "red")

# the to calculate the area of the symmetric difference
st_area(symdiff1)
```

## Hausdorff distance
https://rdrr.io/cran/tigers/man/HausdorffDistance.html
```{r}
library("tigers")

points.og1 = st_cast(sc.og.hs, "MULTIPOINT") %>% as("Spatial") %>% coordinates()
points.sf1 = st_cast(sc.sf.hs, "MULTIPOINT") %>% as("Spatial") %>% coordinates()

# does not give the answer
HausdorffDistance(points.og1, points.sf1, directed = FALSE)
st_distance(sc.og.hs, sc.sf.hs, which = "Hausdorff")


## gives the same answer at least
C = rbind(c(2,1), c(10,2),c(5,2), c(1,2), c(2,1)) %>% list() %>% st_polygon()
D = rbind(c(0,1), c(10,0), c(0,0), c(0,1)) %>% list() %>% st_polygon()


C = rbind(c(0,0), c(4,0), c(0,2), c(0,0)) %>% list() %>% st_polygon()
D = rbind(c(2,2), c(6,4), c(2,4), c(2,2)) %>% list() %>% st_polygon()


CC = st_cast(C, "MULTIPOINT") %>% as("Spatial") %>% coordinates()
DD = st_cast(D, "MULTIPOINT") %>% as("Spatial") %>% coordinates()

st_distance(C, D, which = "Frechet")
HausdorffDistance(CC,DD, directed = FALSE)

sqrt(8)
5.09902^2 # 26
2.121326^2 # 4.500024


library("rgeos")

p1 = readWKT("POLYGON((0 0, 4 0, 0 2, 0 0))")
p2 = readWKT("POLYGON((2 2, 6 4, 2 4, 2 2))")
gDistance(og.hs, sf.hs, hausdorff = TRUE)


library("pracma")
hausdorff_dist(as.matrix(as.data.frame(points.og1)), as.matrix(as.data.frame(points.sf1)))

# https://math.stackexchange.com/questions/5002049/how-to-compute-the-hausdorff-distance-between-two-circles-in-mathbb-r2


library("stampr")
library(sp)
stamp.distance()

SpatialPolygonsDataFrame(list(as.SpatialPolygons.PolygonsList(sc.og.hs), as.SpatialPolygons.PolygonsList(sc.sf.hs)))
is(sc.og.hs)
```
FINAL VERSION
we know this works at least in the right way and it is still widely available so going with this Hausdorff distance from tigers
```{r}
A = rbind(c(0,0), c(4,0), c(0,2))
B = rbind(c(2,2), c(6,4), c(2,4))

s = 1:nrow(A)
rbind(A,B)
D = as.matrix(dist(rbind(A,B)));D
D[s,-s]


HausdorffDistance <- function(A, B, directed = FALSE)
{
    s <- 1:nrow(A)
    D <- as.matrix(dist(rbind(A, B)))
    D <- D[s, -s]
    h <- max(apply(D, 1, min))
    if (directed) return(h)
    max(h, apply(D, 2, min))
}

HausdorffDistance(A,B)
```




## Frechet distance
```{r}
A = st_polygon(list(rbind(c(0,0), c(4,0), c(0,2), c(0,0))))
B = st_polygon(list(rbind(c(2,2), c(6,4), c(2,4), c(2,2))))


st_distance(A, B, which = "Frechet", by_element = TRUE)


points.og1 = st_cast(sc.og.hs, "MULTIPOINT") %>% as("Spatial") %>% coordinates() %>% list() %>% st_polygon()
points.sf1 = st_cast(sc.sf.hs, "MULTIPOINT") %>% as("Spatial") %>% coordinates() %>% list() %>% st_polygon()

st_distance(points.og1, points.sf1, which = "Frechet", by_element = TRUE)

```

```{r}
# Load required libraries
library(ggplot2)
library(gridExtra)

# Define point sets
A <- matrix(c(0, 0,
              4, 0,
              0, 2), ncol = 2, byrow = TRUE)

B <- matrix(c(2, 2,
              6, 4,
              2, 4), ncol = 2, byrow = TRUE)

A = st_cast(sc.og.hs, "MULTIPOINT") %>% as("Spatial") %>% coordinates() %>% as.matrix()
B = st_cast(sc.sf.hs, "MULTIPOINT") %>% as("Spatial") %>% coordinates() %>% as.matrix()



# Euclidean distance function
euclidean <- function(p1, p2) {
  sqrt(sum((p1 - p2)^2))
}

# Compute the discrete Frechet distance matrix
compute_frechet_matrix <- function(P, Q) {
  m <- nrow(P)
  n <- nrow(Q)
  D <- matrix(0, nrow = m, ncol = n)
  
  for (i in 1:m) {
    for (j in 1:n) {
      d <- euclidean(P[i, ], Q[j, ])
      if (i == 1 && j == 1) {
        D[i, j] <- d
      } else if (i == 1) {
        D[i, j] <- max(D[i, j - 1], d)
      } else if (j == 1) {
        D[i, j] <- max(D[i - 1, j], d)
      } else {
        D[i, j] <- max(min(D[i - 1, j], D[i - 1, j - 1], D[i, j - 1]), d)
      }
    }
  }
  return(D)
}

# Compute Frechet distance
frechet_matrix <- compute_frechet_matrix(A, B)
frechet_distance <- frechet_matrix[nrow(A), nrow(B)]
cat(sprintf("Discrete Frechet Distance: %.3f\n", frechet_distance))

# Backtrack to find leash path
trace_path <- function(D) {
  i <- nrow(D)
  j <- ncol(D)
  path <- list(c(i, j))
  
  while (i > 1 || j > 1) {
    if (i > 1 && j > 1) {
      vals <- c(D[i - 1, j], D[i - 1, j - 1], D[i, j - 1])
      min_val <- min(vals)
      if (D[i - 1, j - 1] == min_val) {
        i <- i - 1
        j <- j - 1
      } else if (D[i - 1, j] == min_val) {
        i <- i - 1
      } else {
        j <- j - 1
      }
    } else if (i > 1) {
      i <- i - 1
    } else {
      j <- j - 1
    }
    path <- append(path, list(c(i, j)))
  }
  return(path)
}

# Get the leash path
path <- trace_path(fm)

D = Drotate
C = Crotate

# Prepare data for plotting
df_A <- data.frame(x = C[, 1], y = C[, 2], label = paste0("A", 1:nrow(C)))
df_B <- data.frame(x = D[, 1], y = D[, 2], label = paste0("B", 1:nrow(D)))

# Leash lines
leash_lines <- data.frame()
for (p in path) {
  i <- p[1]
  j <- p[2]
  leash_lines <- rbind(leash_lines, data.frame(
    x = C[i, 1], y = C[i, 2],
    xend = D[j, 1], yend = D[j, 2]
  ))
}

# Plotting
ggplot() +
  geom_path(data = df_A, aes(x = x, y = y), color = "blue") +
  geom_point(data = df_A, aes(x = x, y = y), color = "blue") +
  geom_text(data = df_A, aes(x = x + 0.2, y = y, label = label), color = "blue") +
  geom_path(data = df_B, aes(x = x, y = y), color = "red") +
  geom_point(data = df_B, aes(x = x, y = y), color = "red") +
  geom_text(data = df_B, aes(x = x + 0.2, y = y, label = label), color = "red") +
  geom_segment(data = leash_lines, aes(x = x, y = y, xend = xend, yend = yend), 
               linetype = "dashed", color = "gray40") +
 # ggtitle(sprintf("Discrete Fréchet Distance: %.3f", frechet_distance)) +
  coord_fixed() +
  theme_minimal()

```

```{r}
sqrt(8)
2.5^2
```


#### selecting the right starting points
1. Point closest to the origin (either A or B becomes the first set)
1.1 if multiple options exist pick one closest to the other set
1.2 if still mutliple options exist take the one with the smallest x value
1.3 if still multiple options exist take the one with the smallest y value
2. repeat for the other set, replacing origin with the starting point in the reference set

```{r}
find_closest_pair <- function(A, B) {
  min_dist <- Inf
  min_i <- 1
  min_j <- 1
  for (i in 1:nrow(A)) {
    for (j in 1:nrow(B)) {
      dist <- sqrt(sum((A[i, ] - B[j, ])^2))
      if (dist < min_dist) {
        min_dist <- dist
        min_i <- i
        min_j <- j
      }
    }
  }
  list(i = min_i, j = min_j, distance = min_dist)
}

rotate_curve <- function(curve, start_index) {
  rbind(curve[start_index:nrow(curve), ], curve[-(start_index:nrow(curve)), ], curve[start_index, ])
}

fcp = find_closest_pair(A,B)

A = rotate_curve(A, fcp$i);A
B = rotate_curve(B, fcp$j);B

```

1. pick point closest to the origin
```{r}
# we take sqrt(x^2 + y^2)
# have to think about the fact that we have end = start, so maybe introduce that into the official version, otherhwise always two options

C <- matrix(c(-1,-2,1.5,2,-1,
              -1,2,1.5,-2,-1), ncol = 2, byrow = FALSE)
D <- matrix(2*c(-.75,-2,.5,2,-.75,
                -.75,2,.5,-2,-.75), ncol = 2, byrow = FALSE)

plot(D, col = "blue", type = "b")
lines(C, col  = "red", type = "b")

all_points = cbind(rbind(C,D),group = c(rep(0, nrow(C)), rep(1, nrow(D)))) # incude and indicator from which polygon they are

dist2origin = sqrt(all_points[,1]^2 + all_points[,2]^2)
closest_point = all_points[dist2origin == min(dist2origin),] # all possible closest points to the origin

# now the shortest distance from these points to the points in their other polygon
# so if closest_point[,3] == 1 then ... and == 2 then ...
# for every points in the closest points we select the points that belong to the other polygon and then take minimum distance
fromC = which(closest_point[,3] == 0)
fromD = which(closest_point[,3] == 1)

# for every points we select the rows that have C[,3] == abs(D[,3]-1) and vice versa, and then compare the smallest distance
point1 = closest_point[1,] # just one points as try out
points2comp = all_points[all_points[,3] == abs(point1[3] - 1),] # the points we need to compare to
dist2other = sqrt(all_points[,1]^2 + all_points[,2]^2)


###################

```
```{r}

C <- matrix(c(-1,-2,1.5,2,-1,
              -1,2,1.5,-2,-1), ncol = 2, byrow = FALSE)
D <- matrix(2*c(-.75,-2,.5,2,-.75,
                -.75,2,.5,-2,-.75), ncol = 2, byrow = FALSE)

all_points = as.data.frame(cbind(rbind(C,D),group = c(rep(0, nrow(C)), rep(1, nrow(D))))) # include and indicator from which polygon they are
all_points = all_points[!duplicated(all_points[,1:2]),]
all_points$dist2origin = sqrt(all_points[,1]^2 + all_points[,2]^2)

df1 = all_points[all_points$dist2origin == min(all_points$dist2origin),]

# so now for each points the minimum distance to the other matrix
df1$min_dist =  apply(df1, 1, function(x) {
  opo = all_points[all_points$group == abs(x[3] - 1),]
  min(sqrt((opo$V1 - x[1])^2 + (opo$V2 - x[2])^2))
  
  })   

df1 = df1[df1$min_dist == min(df1$min_dist),]

# now we take the smallest x coordinates (use if statemnet here)
if(nrow(df1) > 1){
  start = df1[which(df1$V1 == min(df1$V1)),] # row with the smallest x values

  if(nrow(start) > 1) start = df1[which(start$V2 == min(start$V2)),]
  
}


mindistance_to_point = function(data, point){
  # note that data has to have data[,1] -> x coord, and data[,2] -> y coord
  data$dist2point = sqrt((data[,1] - point[1])^2 + (data[,2] - point[2])^2)
  data[data$dist2point == min(data$dist2point),]
}



find_starting_point = function(C, D, origin = c(0,0)){
  if(missing(D)) all_points = as.data.frame(C)  # include and indicator from which polygon they are
  else all_points = as.data.frame(cbind(rbind(C,D), group = c(rep(0, nrow(C)), rep(1, nrow(D)))))  # include and indicator from which polygon they are
  print(all_points)
  
  all_points = all_points[!duplicated(all_points[,1:2]),] # remove duplicate points based on x and y (if optimal point occurs in C and D it will be automatically picked later)
  
  # distance from each point to the origin
  df1 = mindistance_to_point(all_points, origin)
  
  # distance to other matrix (if applicable)
  if(origin[1] == 0 & origin[2] == 0) {
    df1$min_dist = apply(df1, 1, function(x) {
      print(x)
      opo = all_points[all_points$group == abs(x[3] - 1),]
      print(opo)
      min(sqrt((opo$V1 - x[1])^2 + (opo$V2 - x[2])^2))
      }) 
    
      df1 = df1[df1$min_dist == min(df1$min_dist),]
  }
  
  # now we take the smallest x coordinates and if they are the same the smallest y value
  if(nrow(df1) > 1){
    start = df1[which(df1$V1 == min(df1$V1)),] # row with the smallest x values
  
    if(nrow(start) > 1) start = df1[which(start$V2 == min(start$V2)),]
  } else start = df1
  
  
  # now find the point on the other polygon
  start2 = find_starting_point(C = all_points[all_points$group == abs(x[3] - 1),], origin = c(start$V1, start$V2))
  
  return(rbind(start, start2))
}

g5 = find_starting_point(C = C,D = D)

g3 = find_starting_point(C,D)
g4 = find_starting_point(C = D,origin = c(g3$V1, g3$V2))

r = NULL
is.numeric(r)

```


```{r}

# Function to calculate Euclidean distance between two points
euclidean_dist <- function(p1, p2) {
  sqrt((p1[1] - p2[1])^2 + (p1[2] - p2[2])^2)
}

# Function to find the point in either C or D closest to the origin
closest_to_origin <- function(C, D) {
  # Calculate distances of each point in C and D from the origin (0, 0)
  origin_distances_C <- apply(C, 1, function(point) euclidean_dist(data.frame(x = 0, y = 0), point))
  origin_distances_D <- apply(D, 1, function(point) euclidean_dist(data.frame(x = 0, y = 0), point))
  
  print(origin_distances_D)
  # Get the minimum distances for each set
  min_dist_C <- min(origin_distances_C)
  min_dist_D <- min(origin_distances_D)
  
  # Determine which set has the point closest to the origin
  if (min_dist_C < min_dist_D) {
    # Choose point from C
    closest_points <- C[origin_distances_C == min_dist_C, ]
    source_set <- "C"
  } else if (min_dist_D < min_dist_C) {
    # Choose point from D
    closest_points <- D[origin_distances_D == min_dist_D, ]
    source_set <- "D"
  } else {
    # If both are equally close, pick from the set with the closest point to the other set
    dist_to_other_set_C <- apply(C, 1, function(point) min(apply(D, 1, function(d) euclidean_dist(point, d))))
    dist_to_other_set_D <- apply(D, 1, function(point) min(apply(C, 1, function(c) euclidean_dist(point, c))))
    
    if (min(dist_to_other_set_C) < min(dist_to_other_set_D)) {
      # Pick point from C
      closest_points <- C[origin_distances_C == min_dist_C, ]
      source_set <- "C"
    } else {
      # Pick point from D
      closest_points <- D[origin_distances_D == min_dist_D, ]
      source_set <- "D"
    }
  }
  print(closest_points)
  
  # If there are multiple candidates, resolve tie by choosing the smallest x and y
  closest_points <- closest_points[order(closest_points$x, closest_points$y), ]
  return(list(point = closest_points[1, ], source_set = source_set))
}

# Function to find the point in the other set (D or C) closest to the chosen point
closest_to_chosen_point <- function(chosen_point, source_set, C, D) {
  # If the chosen point was from C, find the closest point in D, and vice versa
  if (source_set == "C") {
    dist_to_D <- apply(D, 1, function(point) euclidean_dist(chosen_point, point))
    closest_points <- D[dist_to_D == min(dist_to_D), ]
  } else {
    dist_to_C <- apply(C, 1, function(point) euclidean_dist(chosen_point, point))
    closest_points <- C[dist_to_C == min(dist_to_C), ]
  }
  
  # Resolve tie by choosing the smallest x and y
  print(closest_pointx)
  closest_points <- closest_points[order(closest_points$x, closest_points$y), ]
  return(closest_points[1, ])
}

# Step 1: Find the first point closest to the origin (it could be from C or D)
chosen_first <- closest_to_origin(C, D)

# Step 2: Find the closest point from the other set to the chosen first point
chosen_second <- closest_to_chosen_point(chosen_first$point, chosen_first$source_set, C, D)

# Print the results
cat("Chosen first point (from set", chosen_first$source_set, "):\n")
print(chosen_first$point)
cat("Chosen second point (from the other set):\n")
print(chosen_second)

```


```{r}

all_points = as.data.frame(cbind(rbind(C,D), group = as.factor(c(rep("C", nrow(C)), rep("D", nrow(D)))))) # include and indicator from which polygon they are
all_points = all_points[!duplicated(all_points[,1:2]),]

# IF statement so we can reuse for finding the second starting point 

all_points$dist2origin = sqrt(all_points[,1]^2 + all_points[,2]^2)

# find the points that are closest to the origin (could be multiple from the same group and from different groups)
closest2origin = all_points %>% filter(dist2origin == min(dist2origin)) 

# IF statement if the if we have more than one row

# for every point calculate how close it is to the other group
closest2origin$dist2other = apply(closest2origin, 1, function(x) {
  other_group = all_points[all_points$group != as.numeric(x[3]),]
  min(sqrt((x[1] - other_group[,1])^2 + (x[2]- other_group[,1])^2))
})

closest2other = closest2origin %>% filter(dist2other == min(dist2other)) 

# IF we have more than one row

# pick points which has the lowest x value
smallest.x = closest2other[which(closest2other$V1 == min(closest2other$V1)), ]

# IF still more than one row

# pick points with lowest y value
smallest.y = closest2other[which(smallest.x$V2 == min(smallest.x$V2)), ]

```

```{r}
first_starting_point = function(C, D, first = c(0,0)){
    
  all_points = as.data.frame(cbind(rbind(C,D), group = as.factor(c(rep("C", nrow(C)), rep("D", nrow(D)))))) # include and indicator from which polygon they are
  all_points = all_points[!duplicated(all_points[,1:2]),]
    
  # find distance to origin 
  all_points$dist2origin = sqrt((all_points[,1] - first[1])^2 + (all_points[,2] - first[2])^2)
  
  # find the points that are closest to the origin (could be multiple from the same group and from different groups)
  closest2origin = all_points %>% filter(dist2origin == min(dist2origin)) 

  # if we have more than one row we calculate the minimal distance of those points to the other group
  if (nrow(closest2origin) == 1) start1 = closest2origin 
  else {
    closest2origin$dist2other = apply(closest2origin, 1, function(x) {
      other_group = all_points[all_points$group != as.numeric(x[3]),] # define other group
      min(sqrt((x[1] - other_group[,1])^2 + (x[2]- other_group[,2])^2))}) # minimal distance to other group
    
    closest2other = closest2origin %>% filter(dist2other == min(dist2other)) 

    if (nrow(closest2other) == 1) start1 = closest2other 
    else {
      smallest.x = closest2other[which(closest2other$V1 == min(closest2other$V1)), ]
      
      if(nrow(smallest.x) == 1) start1 = smallest.x 
      else start1 = closest2other[which(smallest.x$V2 == min(smallest.x$V2)), ]
    }
  }
  
  return(start1)
  
}

hs1 = first_starting_point(C,D)
hs1[1:2]

sqrt(.5^2)

```

```{r}
second_starting_point = function(C,D,start) {
  
  all_points = as.data.frame(cbind(rbind(C,D), group = as.factor(c(rep("C", nrow(C)), rep("D", nrow(D)))))) # include and indicator from which polygon they are
  all_points = all_points[!duplicated(all_points[,1:2]),]
  relevant_points = all_points %>% filter(group != start$group)

  # if the first starting point also occurs in the other group this is automatically the second starting point
  if(nrow(merge(data.frame(start[,1:2]), relevant_points[,1:2]))>0) start2 = start
  else {
    relevant_points$dist2start = sqrt((relevant_points[,1] - start[1,1])^2 + (relevant_points[,2] - start[1,2])^2)
    closest2start = relevant_points %>% filter(dist2start == min(dist2start)) 
    
    if (nrow(closest2start) == 1) start2 = closest2start 
    else {
      closest2start$dist2other = apply(closest2start, 1, function(x) {
        other_group = all_points[all_points$group != as.numeric(x[3]),] # define other group
        min(sqrt((x[1] - other_group[,1])^2 + (x[2]- other_group[,2])^2))}) # minimal distance to other group
      
      closest2other = closest2start %>% filter(dist2other == min(dist2other)) 
  
      if (nrow(closest2other) == 1) start2 = closest2other 
      else {
        smallest.x = closest2other[which(closest2other$V1 == min(closest2other$V1)), ]
        
        if(nrow(smallest.x) == 1) start2 = smallest.x 
        else start2 = closest2other[which(smallest.x$V2 == min(smallest.x$V2)), ]
      }
    }    
    
  }
  
  return(start2)
}

second_starting_point(C,D,hs1)




```

```{r}

C <- matrix(c(-1,-2,1.5,2,-1,
              -1,2,1.5,-2,-1), ncol = 2, byrow = FALSE)
D <- matrix(2*c(-0.5,-0.75,-2,.5,2,-.5,
                -0.75,-0.5,2,.5,-2,-.75), ncol = 2, byrow = FALSE)

plot(D, col = "blue", type = "b", asp = 1)
lines(C, col  = "red", type = "b")
```







##### now also have to make sure that we go clockwise within the points
```{r}
plot(0:5, asp = 1, type = "n")
polygon(A, col = "red"); points(A, lwd = 5, col = "pink")
polygon(B, col = "blue"); points(B, lwd = 5, col = "lightblue")
```

```{r}
Ap = st_polygon(list(as.matrix(A[,1:2])))
Bp = st_polygon(list(as.matrix(B[,1:2])))

AC = st_centroid(Ap)
BC = st_centroid(Bp)

AC[2] # y and [1] is x

a = atan( (-2) / (-2) );a * (180/pi)

angle_func = function(c, p){
  atan( (c[2] - p[2]) / (c[1] - p[1])) * ( (180 / pi))
}

a = c(0,0) # note that since all of our centrouds are centered we will always compare to 0,0,  not all measures describe this tho
b = c(2,-2)
angle_func(a, b)


atan2(b[2], b[1]) * ( (180 / pi))

apply(A,1,function(x) angle_func(x,AC))

C <- matrix(c(-1, 1,
              2,2,
              2,-2,
              -1,-1), ncol = 2, byrow = TRUE)

angle_func2 = function(p){
  degrees = atan2(p[2], p[1]) * ( (180 / pi))
  #ifelse(degrees < 0, degrees + 360, degrees)
}

apply(C,1,angle_func2)

```

```{r}
# Example polygon (irregular shape)
points <- data.frame(
  x = c(1,-1,1,3,3,1),
  y = c(1,0,2,2,0,-1)
)

plot(points, type = "b")

# Compute centroid (mean method, works for many shapes)
centroid <- colMeans(points)

# Compute angle from centroid to each point
angles <- atan2(points$y - centroid[2], points$x - centroid[1])
angles_deg <- angles * 180 / pi
angles_deg[angles_deg < 0] <- angles_deg[angles_deg < 0] + 360

# Order clockwise (descending)
cw_order <- order(-angles_deg)
points_cw <- points[cw_order, ]

# Plot to verify
plot(points, type = 'n', asp = 1, main = "Clockwise Polygon")
polygon(points_cw, col = 'lightblue', border = 'blue')
points(centroid[1], centroid[2], col = "red", pch = 19)
text(points_cw, labels = seq_len(nrow(points_cw)), pos = 3)

```
```{r}
# Load necessary packages
library(ggplot2)
library(dplyr)

# Define a crescent-like, non-convex polygon (deliberately weird shape)
crescent <- data.frame(
  x = c(3,2,4, 5, 4, 2, 1, 0, 1, 3) ,#%>% rev(),
  y = c(1,0,0, 2, 4, 4, 3, 2, 1, 1) #%>% rev()
)

plot(crescent, type = "l")

# Close the polygon for visualization
crescent_closed <- rbind(crescent, crescent[1, ])

# Function to check winding direction
is_clockwise <- function(x, y) {
  n <- length(x)
  x_next <- c(x[-1], x[1])
  y_next <- c(y[-1], y[1])
  area <- sum(x * y_next - x_next * y)
  print(area)
  return(area < 0)
}

# Check direction
cw <- is_clockwise(crescent$x, crescent$y)
cat("Original winding: ", ifelse(cw, "Clockwise", "Counter-Clockwise"), "\n")

# Optionally reverse the polygon if you want clockwise order
if (!cw) {
  crescent <- crescent[nrow(crescent):1, ]
}
crescent_closed <- rbind(crescent, crescent[1, ])

# Add index for arrow labels
crescent$idx <- 1:nrow(crescent)

# Plot
ggplot(crescent, aes(x, y)) +
  geom_polygon(fill = "lightblue", color = "blue", alpha = 0.5) +
  geom_path(arrow = arrow(type = "open", length = unit(0.2, "inches")),
            color = "black", size = 1) +
  geom_point(size = 3, color = "red") +
  geom_text(aes(label = idx), vjust = -1, fontface = "bold") +
  coord_fixed() +
  ggtitle(paste("Polygon with", ifelse(cw, "Clockwise", "Counter-Clockwise"), "Winding")) +
  theme_minimal()


is_clockwise <- function(x, y) {
  n <- length(x)
  x_next <- c(x[-1], x[1])
  y_next <- c(y[-1], y[1])
  area <- sum(x * y_next - x_next * y)
  print(area)
  return(area < 0)
}

make_clockwise = function(coordinates){
  # check direction and change if needed
  direction = is_clockwise(coordinates[,1], coordinates[,2])
  
  if(!direction) return(coordinates[nrow(coordinates):1, ])
  else return(coordinates)
  
}



```

FINAL FRECHET DISTANCE EVERYTHING COMBINED
```{r}
## first we calculate the starting points
###### first starting points
first_starting_point = function(C, D, first = c(0,0)){
    
  all_points = as.data.frame(cbind(rbind(C,D), group = as.factor(c(rep("C", nrow(C)), rep("D", nrow(D)))))) # include and indicator from which polygon they are
  all_points = all_points[!duplicated(all_points[,1:2]),]
    
  # find distance to origin 
  all_points$dist2origin = sqrt((all_points[,1] - first[1])^2 + (all_points[,2] - first[2])^2)
  
  # find the points that are closest to the origin (could be multiple from the same group and from different groups)
  closest2origin = all_points %>% filter(dist2origin == min(dist2origin)) 

  # if we have more than one row we calculate the minimal distance of those points to the other group
  if (nrow(closest2origin) == 1) start1 = closest2origin 
  else {
    closest2origin$dist2other = apply(closest2origin, 1, function(x) {
      other_group = all_points[all_points$group != as.numeric(x[3]),] # define other group
      min(sqrt((x[1] - other_group[,1])^2 + (x[2]- other_group[,2])^2))}) # minimal distance to other group
    
    closest2other = closest2origin %>% filter(dist2other == min(dist2other)) 

    if (nrow(closest2other) == 1) start1 = closest2other 
    else {
      smallest.x = closest2other[which(closest2other$V1 == min(closest2other$V1)), ]
      
      if(nrow(smallest.x) == 1) start1 = smallest.x 
      else start1 = closest2other[which(smallest.x$V2 == min(smallest.x$V2)), ]
    }
  }
  
  return(start1)
  
}

##### second starting points
second_starting_point = function(C,D,start) {
  
  all_points = as.data.frame(cbind(rbind(C,D), group = as.factor(c(rep("C", nrow(C)), rep("D", nrow(D)))))) # include and indicator from which polygon they are
  all_points = all_points[!duplicated(all_points[,1:2]),]
  relevant_points = all_points %>% filter(group != start$group)

  # if the first starting point also occurs in the other group this is automatically the second starting point
  if(nrow(merge(data.frame(start[,1:2]), relevant_points[,1:2]))>0) start2 = start
  else {
    relevant_points$dist2start = sqrt((relevant_points[,1] - start[1,1])^2 + (relevant_points[,2] - start[1,2])^2)
    closest2start = relevant_points %>% filter(dist2start == min(dist2start)) 
    
    if (nrow(closest2start) == 1) start2 = closest2start 
    else {
      closest2start$dist2other = apply(closest2start, 1, function(x) {
        other_group = all_points[all_points$group != as.numeric(x[3]),] # define other group
        min(sqrt((x[1] - other_group[,1])^2 + (x[2]- other_group[,2])^2))}) # minimal distance to other group
      
      closest2other = closest2start %>% filter(dist2other == min(dist2other)) 
  
      if (nrow(closest2other) == 1) start2 = closest2other 
      else {
        smallest.x = closest2other[which(closest2other$V1 == min(closest2other$V1)), ]
        
        if(nrow(smallest.x) == 1) start2 = smallest.x 
        else start2 = closest2other[which(smallest.x$V2 == min(smallest.x$V2)), ]
      }
    }    
    
  }
  
  return(start2)
}


#### rotate to clockwise if necessary
is_clockwise <- function(x, y) {
  n <- length(x)
  x_next <- c(x[-1], x[1])
  y_next <- c(y[-1], y[1])
  area <- sum(x * y_next - x_next * y)
  return(area < 0)
}

make_clockwise = function(coordinates){
  # check direction and change if needed
  direction = is_clockwise(coordinates[,1], coordinates[,2])
  
  if(!direction) return(coordinates[nrow(coordinates):1, ])
  else return(coordinates)
  
}


### actual calculating of the frechet distance
compute_frechet_matrix <- function(P, Q) {
  m <- nrow(P)
  n <- nrow(Q)
  D <- matrix(0, nrow = m, ncol = n)
  
  for (i in 1:m) {
    for (j in 1:n) {
      d <- euclidean(P[i, ], Q[j, ])
      if (i == 1 && j == 1) {
        D[i, j] <- d
      } else if (i == 1) {
        D[i, j] <- max(D[i, j - 1], d)
      } else if (j == 1) {
        D[i, j] <- max(D[i - 1, j], d)
      } else {
        D[i, j] <- max(min(D[i - 1, j], D[i - 1, j - 1], D[i, j - 1]), d)
      }
    }
  }
  
  return(D)
  
}

```


```{r}
## the data
C <- matrix(c(-1,-2,1.5,2,-1,
              -1,2,1.5,-2,-1), ncol = 2, byrow = FALSE)

C <- matrix(c((c(-1,-2,1.5,2,-1) %>% rev()),(c(-1,2,1.5,-2,-1) %>% rev())), ncol = 2, byrow = FALSE)
D <- matrix(2*c(-0.5,-0.75,-2,.5,2,-.5,
                -0.75,-0.5,2,.5,-2,-.75), ncol = 2, byrow = FALSE)

## deduplicate the data
C = C[2:nrow(C),]
D = D[2:nrow(D),]

firsths = first_starting_point(C,D)
secondhs = second_starting_point(C,D,firsths)

Cstart = which(C[,1] == firsths[,1] & C[,2] == firsths[,2])
Dstart = which(D[,1] == secondhs[,1] & D[,2] == secondhs[,2])

Corder = rbind(C[Cstart:nrow(C),], C[1:Cstart,])
Dorder = rbind(D[Dstart:nrow(D),], D[1:Dstart,])

Crotate = make_clockwise(Corder)
Drotate = make_clockwise(Dorder)

fm = compute_frechet_matrix(Crotate, Drotate)
frechet_distance <- fm[nrow(Crotate), nrow(Drotate)]


calc_frechet_distance = function(C,D){
  # remove duplicates
  C = C[!duplicated(C),]
  D = D[!duplicated(D),]
  
  firsths = first_starting_point(C,D)
  secondhs = second_starting_point(C,D,firsths)
  
  Cstart = which(C[,1] == firsths[,1] & C[,2] == firsths[,2])
  Dstart = which(D[,1] == secondhs[,1] & D[,2] == secondhs[,2])
  
  Corder = rbind(C[Cstart:nrow(C),], C[1:Cstart,])
  Dorder = rbind(D[Dstart:nrow(D),], D[1:Dstart,])
  
  Crotate = make_clockwise(Corder)
  Drotate = make_clockwise(Dorder)
  
  fm = compute_frechet_matrix(Crotate, Drotate)
  frechet_distance <- fm[nrow(C), nrow(D)]
  
  return(frechet_distance)
}

calc_frechet_distance(C,D)



```


group D with x = (-1.5,-4,1,4,-1,-1.5) and y = c(-1,4,1,-4,-1.5,-1)
group C with x = (-1,-2,1.5,2,-1) and y = c(-1,2,1.5,-2,-1)


```{r}
A = st_polygon(list(Crotate))
B = st_polygon(list(Drotate))

plot(A)
plot(B, add = TRUE)

st_distance(A, B, which = "Frechet", by_element = TRUE)
st_distance()
```











## box-counting dimension



































